Delta Factor - A New Authorization Factor

### 2022-08-09 paper1.tex
submitted

%---------------------------------------------------------------------------------------------------------------------
% Following exactly professor's Canvas example
\section{Abstract}
\setlength{\baselineskip}{1.5\baselineskip}
((TBA as per professor - however...))\\ %This is modus operandi bc as the paper evolves, the abstract changes invariably
\noindent
Man\_In\_The\_Middle-ware, (MITM) is a rapidly growing field of website spoofing that is a challenge for everyone, especially for MITM mitigation and deterrence  developers to keep up with. Many papers have been published that explain the nuances, methods and lengths that the attackers use for "personal gain". According to [1], fairly spectacular results in detecting a MITM attack has been shown. This was accomplished by the group gathering attack data, metrics of the attacks, and analysis for many months, to eventually develop a custom phishing toolkit they named "phoca". That said, the target of our research is two-fold. Firstly, we show improvement in the Stonybrook claim via our novel composite-type methodology using COTP.  Secondly, we claim and show that despite Stonybrook's claim, it is possible and provable that application level checks based on easily-acccessable and well known network level features along with a secret, can indeed determine with fine granularity, a MITM or !MITM. Ie., an efficient and novel server authentication factor has been developed.

@@

There are methods to avoid and/or determine the existence of a MITM. Traditionally there are eg., Machine Learning methods, ie., the ubiquitous trust-based methods, 2FA, and more new ways, methods, and tools for detection of MITM and more, however they aren't used nor easily available to the general public.
The current status in the field, including the best performing tool is from the basis of our research, with is from the SUNY - Stonybrook paper [1], "Catching Transparent Phish: Analyzing and Detecting MITM  Phishing Toolkits". The name of the extremely efficient tool that they developed is called, "phoca". In brief however, it is discovered that via COTP and of course a shared secret plus a 2FA for security, the APPLICATION layer CAN BE able to help us reveal a MITM if one exists, as opposed to the denial of that concept by Stonybrook, again, knowing that they did a tremendous amount of work on their phoca toolkit.

@@

\noindent
The relevance and importance of COTP will also be explained.\\

@@

Although these works are focused on finding transparent phishing sites, or detection of the attack from an firewall or hardware perspective - blocking the client from access the site prior to the attack taking place. These papers undersline the importance of utilizing Network-level / connection features to detect advanced MITMs. Key difference between those and this work is fundamentally looking at client and server interactions, and attempting to detect a MITM. Ultimately empowering both client and server to determine for if an MITM is present.

@@

These works are public standards used for common multi factor authenication methods. Both lean into research accelration of cryptographic functions and the need for client validation. What seperates these from this work, is that these methods are fundamentally weak against an advanced MITM attack. Even in the scenario where both client and server validation is in place, an advanced MITM can be undetected. This work attempts to enhance the prior methods by drawing from shared connection features.

@@

\subsection{One-Time Passwords}
Earlier web servers only used knowledge to validate or authenicate users. But modern applications employ multi factor approaches to make a more robust authenication method. The most basic of these is the HMAC-based One-Time Password (HOTP). Further factors are introduced, such as time, creating additional constraints on authenication. These methods rely on cryptographic functions that are fast computations in one direction, while computationally complex to reverse.

These passwords are made by the server generating a high entropy secret. This secret is shared with the client over a secure channel. Then both client and server, use cryptographic functions to convert this shared secret into a short password (sometimes using additional factors outside of the shared secret). The client then enteres the generated password, and the server compares it's generated version to the client provided password. If they match, the server allows the client to access protected resources that the client should have access too. Otherwise, the server rejects the client, denying access to the protected resources.

@@

% Paragraph 2,3,...: introduction of DT techniques ... [Actually - Inspiration]
Inspired by an existing movement, a growing trend towards multifactor authentication
identified that most methods still only do knowledge-based or device-based validation. This is by incorporating existing and standardized One-Time-Pass code technology, with connection-oriented details (eg., COTP) to produce a method to offer an authentication much stronger than knowledge-based, possesion-based, or time-based alone.

\subsection{Webserver}
Webservers have been the focus on security, just as Web Browsers have been. Servers also have the most access, as they usually the trusted resource by default, allowing for phishing attacks to happen in the first place. They hold protected resources, and have implied trust. However, this paper suggests that Webservers should also offer to build trust with the client - instead of being implictly trusted - as is the norm today - but offer to allow client to validate the connection, knowledge, time, through an One-Time Password as well. This means that the server will accept a shared secret, just like a client would for a multi factor authenication. Standard modules are built into NGINX to extract TLS and TCP connection data. They then pass the details onto a Python WebApp for further validation.

\subsection{Web Broswer}
Web browsers the the interface of clients for many resources on the internet. These also have been hardened over the years to prevent leaking of private information. For this reason they are often locked down, where even the client can't directly influence much of the behavior - this is the same access an attacker would have on a compromised system. To get access to the life cycle of requests and network level details, browsers offer enhanced APIs to extension/plug-in developers. This is what we'll use in order to extract the features needed for COTP and calculate it.

@@

The setup for this experiment is a simple proof of concept webserver. The webserver uses URIs to indicate protected data. The client will use a web browser to access this webserver and it's protected resource. The resource will simply be the client's username - and accessing this page will trigger the server generate an one-time password, and embed this information into presented page. The client's web browser also includes the instrumentation to generate a one-time password via a browser extension/plug in.

Normally the experiment would be that some group of users get a traditional 2FA code (either HOTP or TOTP). Another group would be presented with a COTP code instead. Finally a fraction of both groups would have a 'fake' MITM attack. Due to time constaints we'll evaulate the method's contribution to existing traditional methods instead.

@@

This round-trip-time was added as an additional factor in addition to the TLS features described. The same experiment was performned, a client through a web browser would attempt to load a protected resource 20 times, and observed how many times the code was incorrect. In addition, for this experiment, TCP round trip times were also recorded for both client and server.

@@

\subsection{Results + Data}
For the TLS features along, every trial had server and client COTPs match! 
For the TLS + uncorrected TCP features, 2 of 20 trials matched!
For the TLS + corrected TCP features, 16 of 20 trials matched!

\subsubsection{Round Trip Times}
Server RTT ms, Client RTT ms, ABS Diff
21	22	1
21	25	4
18	22	4
19	23	4
19	21	2
19	22	3
18	23	5
21	23	2
18	21	3
17	20	3
19	20	1
18	21	3
19	23	4
19	20	1
20	21	1
18	21	3
18	22	4
19	20	1
20	24	4
19	20	1
19	20	1
18	21	3

\subsubsection{Analysis}
Although TLS features only gave great reproducibility in regards to COTPs it doesn't add very much to the strength of the algorithm. An advanced MITM would only need to observe the TLS hellos to brute force the selected cipher suite, and produce their own COTPs. The 30 second time frame can be reduced to make this less likely; however this isn't the strongest solution.

Combining TLS and TCP features makes the attacker not only brute force the ciphersuite (and the shared secret, and within the timeframe), but they must now also match connection latencies. The MITM needs to literally be in the middle, in terms of latency to both server and client. This is much more difficult, as they could be closer to one party over the other and must compensate. Due to connection information being unencrypted and public, it's not impossible for a motivated attacker to find hosts equidistant - but requires much more effort.

Choosing a window that doesn't typically happen, and that is large enough that small differences don't end up in different timing windows is critical. This work uses a prime number of 7ms to generate windows. Although other primes could be used, 7ms was the sweet spot for this experiment. Note: a similiar approach to generating windows in TOTP is done, a integer division is used to determine the window a particular latency belongs too. (See code for reference)

@@

@@

%---------------------------------------------------------------------------------------------------------------------
\section{Conclusion}
In this work, we looked at providing clients or servers a method of detecting transparent MITMs, one of the most advanced phishing attacks today by using connection details with a proposed algorithm based on one time passwords, called Connection-Oriented One-Time Passwords. Utilizing COTP as an additional factor, once the user has been authenicated (by means of knowledge, or some other factor), COTP enables MITM detection through generation of a composite secret from multiple factors including connection features observed. In other words, instead of the server validating the user again, with this factor, we propose the inverse: the server offers the user an enhanced one time use password to validate. We evaluated some potential weaknesses of the method, and techiques we used to make COTP an useful validation factor, along with a reference implementation using publically available browsers and web server implementations.

\subsubsection{Should not be a Solo Factor}
COTP provides a robust method to validate multiple factors, namely: knowledge, connection details, and device. However, for maximum effectiviness, validation of both parties should be done. First, the user should be validated by the server - ideally using COTP or another 2FA method. This is the same as the state of the art today by Duo or Google Authenticator. But, an additional validation should be used to allow the client to validate the server. Because the server has already validated the client, this is effectively a courtesy offered, and has no impact on authorization to access the protected resources.

\subsubsection{Connection Latency Reporting May Vary}
This work offers a method to deal with latency mismatches between the server and the client by rounding to the nearest prime based latency window. Multiple COTPs should be done through the client, or, simply retrying authentication as proposed in this work. However, this method means that a stochastic method to validating the user should be used to reduce the number of retries. This is in contrast to traditional one-time password methods, which can be performed from a seperate device.

\subsection{Future Work}
Future work includes further addressing timing discrepencies between server and client. One interesting approach could be remap the values to nearest prime number or calibrating the error out during server boot. Care must be taken to ensure that close values map to a simliar value while values with large differences map to entirely different values - all without the counterpart's number. If such a techinque is found, this factor becomes significantly more robust to one time code differences.

Another potiential future work is finding the optimal balance between connection oriented details and other factors. This work focused on introducing connection oriented features to a verification algorithm. This is by using only shared features between both parties, however there are additional features that are shared, which can be included. One such additional feature is each party's supported cipher suite, or, another is the selected curve - and both could increase difficultly for an advanced transparent MITM to sucessfully mimick all details symetrically. However, the increase in difficultly also needs to be assessed. Another issue is balance between authentication features, ie., a completely connection-oriented approach will fail, due to connection details being publically available to a motivated attacker through protocol analysis.

Finally, another future work is potientially repeatedly validating or checking factors periodically. TLS addresses forward secrecy through rekeying, and this happens to be part of the protocol. The same process could trigger a transparent COTP generation and exchange. This type of feature is already possible on the browser side, and could be implemented on the server side as well. In addition, a new shared secret could be generated and shared upon rekeying, forward authentication is allowed. Those with access to previous details would not have future authentication.

>>>>

%---------------------------------------------------------------------------------------------------------------------
% Following exactly professor's Canvas example
\section{Abstract}
\setlength{\baselineskip}{1.5\baselineskip}
%This is modus operandi bc as the paper evolves, the abstract changes invariably
\noindent
Man\_In\_The\_Middle-ware, (MITM)[14],[20] is a rapidly growing field of website spoofing that is a challenge for everyone, especially for MITM mitigation and deterrence  developers to keep up with. Many papers have been published that explain the nuances, methods and lengths that the attackers use for "personal gain". According to [1], fairly spectacular results in detecting a MITM attack has been shown. This was accomplished by the group gathering attack data, metrics of the attacks, and analysis for many months, to eventually develop a custom phishing toolkit they named "phoca". That said, the target of our research is two-fold. Firstly, we show improvement in the Stonybrook claim via our novel composite-type methodology using COTP[13].  Secondly, we claim and show that despite Stonybrook's claim, it is possible and provable that application level checks based on easily acccessible and well known network level features along with a secret, can indeed determine with fine granularity, a MITM or !MITM. Ie., an efficient and novel server authentication factor has been developed.

@@

There are methods to avoid and/or determine the existence of a MITM. Traditionally there are eg., Machine Learning methods, ie., the ubiquitous trust-based methods, 2FA, and more new ways, methods, and tools for detection of MITM and more, however they aren't used nor easily available to the general public.
The current status in the field, including the best performing tool is from the basis of our research, with is from the SUNY - Stonybrook paper [1], "Catching Transparent Phish: Analyzing and Detecting MITM  Phishing Toolkits". The name of the extremely efficient tool that they developed is called, "phoca". In brief however, it is discovered that via COTP[13] and of course a shared secret plus a 2FA for security, the APPLICATION layer CAN BE able to help us reveal a MITM if one exists, as opposed to the denial of that concept by Stonybrook, again, knowing that they did a tremendous amount of work on their phoca toolkit.

@@

\noindent
The relevance and importance of COTP[13] will also be explained.\\

@@

Although these works are focused on finding transparent phishing sites, or detection of the attack from a firewall or hardware perspective - blocking the client from access the site prior to the attack taking place. These papers underline the importance of utilizing Network-level / connection features to detect advanced MITMs. Key difference between those and this work is fundamentally looking at client and server interactions, and attempting to detect a MITM. Ultimately empowering both client and server to determine for if an MITM is present.

@@

These works are public standards used for common multi factor authentication methods. Both lean into research acceleration of cryptographic functions and the need for client validation. What separates these from this work, is that these methods are fundamentally weak against an advanced MITM attack. Even in the scenario where both client and server validation is in place, an advanced MITM can be undetected. This work attempts to enhance the prior methods by drawing from shared connection features.

@@

\subsection{One-Time Passwords}
Earlier web servers only used knowledge to validate or authenticate users. But modern applications employ multi factor approaches to make a more robust authentication method. The most basic of these is the HMAC-based One-Time Password (HOTP)[2][8]. Further factors are introduced, such as time, creating additional constraints on authentication. These methods rely on cryptographic functions that are fast computations in one direction, while computationally complex to reverse.

These passwords are made by the server generating a high entropy[4] secret. This secret is shared with the client over a secure channel. Then both client and server, use cryptographic functions to convert this shared secret into a short password (sometimes using additional factors outside of the shared secret). The client then enters the generated password, and the server compares its generated version to the client provided password. If they match, the server allows the client to access protected resources that the client should have access to. Otherwise, the server rejects the client, denying access to the protected resources.

@@

% Paragraph 2,3,...: introduction of DT techniques ... [Actually - Inspiration]
Inspired by an existing movement, a growing trend towards multi-factor authentication
identified that most methods still only do knowledge-based or device-based validation. This is by incorporating existing and standardized One-Time-Pass code technology, with connection-oriented details (eg., COTP) to produce a method to offer an authentication much stronger than knowledge-based, possession-based, or time-based alone.

\subsection{Webserver}
Webservers have been the focus on security, just as Web Browsers have been. Servers also have the most access, as they usually the trusted resource by default, allowing for phishing attacks to happen in the first place. They hold protected resources, and have implied trust. However, this paper suggests that Webservers should also offer to build trust with the client - instead of being implicitly trusted - as is the norm today - but offer to allow client to validate the connection, knowledge, time, through an One-Time Password as well. This means that the server will accept a shared secret, just like a client would for a multi factor authentication. Standard modules are built into NGINX to extract TLS and TCP connection data. They then pass the details onto a Python WebApp for further validation.

\subsection{Web Browser}
Web browsers the the interface of clients for many resources on the internet. These also have been hardened over the years to prevent leaking of private information. For this reason they are often locked down, where even the client can't directly influence much of the behavior - this is the same access an attacker would have on a compromised system. To get access to the life cycle of requests and network level details, browsers offer enhanced APIs [11],[12],[13] to extension/plug-in developers. This is what we'll use in order to extract the features needed for COTP and calculate it.

@@

The setup for this experiment is a simple proof of concept webserver. The webserver uses URIs to indicate protected data. The client will use a web browser to access this webserver and its protected resource. The resource will simply be the client's username - and accessing this page will trigger the server generate a one-time password, and embed this information into presented page. The client's web browser also includes the instrumentation to generate a one-time password via a browser extension/plug in.

Normally the experiment would be that some group of users get a traditional 2FA code (either HOTP[2],[8] or TOTP[3],[7]). Another group would be presented with a COTP code instead. Finally a fraction of both groups would have a 'fake' MITM attack. Due to time constraints, we'll evaluate the method's contribution to existing traditional methods instead.

@@

This round-trip-time was added as an additional factor in addition to the TLS features described. The same experiment was performed, a client through a web browser would attempt to load a protected resource 20 times, and observed how many times the code was incorrect. In addition, for this experiment, TCP round trip times were also recorded for both client and server.

@@

\subsection{Results + Data}
\noindent 
Please refer to \textbf{Table-1} below.\\
For the TLS features along, every trial had server and client COTPs match! 
For the TLS + uncorrected TCP features, 2 of 20 trials matched!
For the TLS + corrected TCP features, 16 of 20 trials matched!
Please refer to [6]

\subsubsection{\textbf{Table-1} Round Trip Times}

% \includegraphics[width=\textwidth]{pix3}
\includegraphics[width=5cm]{pix3}

% Due to time constraints, and deprecated examples online, I'l use an Excel png		
% Server RTT ms, Client RTT ms, ABS Diff
% 21	22	1
% 21	25	4

% 18	22	4
% 19	23	4
%19	21	2
% 19	22	3
% 18	23	5
% 21	23	2
% 18	21	3
% 17	20	3
% 19	20	1
% 18	21	3

% 19	23	4
% 19	20	1
% 20	21	1
% 18	21	3
% 18	22	4
% 19	20	1
% 20	24	4
% 19	20	1
% 19	20	1
% 18	21	3

\subsubsection{Analysis}
Although TLS features only gave great reproducibility in regards to COTPs [13] it doesn't add very much to the strength of the algorithm. An advanced MITM would only need to observe the TLS hellos to brute force the selected cipher suite, and produce their own COTPs. The 30 second time frame can be reduced to make this less likely; however this isn't the strongest solution.

Combining TLS and TCP features makes the attacker not only brute force the cipher-suite (and the shared secret, and within the timeframe), but they must now also match connection latencies. The MITM needs to literally be in the middle, in terms of latency to both server and client. This is much more difficult, as they could be closer to one party over the other and must compensate. Due to connection information being unencrypted and public, it's not impossible for a motivated attacker to find hosts equidistant - but requires much more effort.

Choosing a window that doesn't typically happen, and that is large enough that small differences don't end up in different timing windows is critical. This work uses a prime number of 7ms to generate windows. Although other primes could be used, 7ms was the sweet spot for this experiment. Note: a similar approach to generating windows in TOTP is done, a integer division is used to determine the window a particular latency belongs too. (See code for reference)

@@

%---------------------------------------------------------------------------------------------------------------------
\section{Conclusion}
In this work, we looked at providing clients or servers a method of detecting transparent MITMs, one of the most advanced phishing attacks today by using connection details with a proposed algorithm based on one time passwords, called Connection-Oriented One-Time Passwords. Utilizing COTP as an additional factor, once the user has been authenticated (by means of knowledge, or some other factor), COTP enables MITM detection through generation of a composite secret from multiple factors including connection features observed. In other words, instead of the server validating the user again, with this factor, we propose the inverse: the server offers the user an enhanced one time use password to validate. We evaluated some potential weaknesses of the method, and techniques we used to make COTP a useful validation factor, along with a reference implementation using publicly available browsers and web server implementations.

\subsubsection{Should not be a Solo Factor}
COTP provides a robust method to validate multiple factors, namely: knowledge, connection details, and device. However, for maximum effectiveness, validation of both parties should be done. First, the user should be validated by the server - ideally using COTP or another 2FA method. This is the same as the state of the art today by Duo or Google Authenticator. But, an additional validation should be used to allow the client to validate the server. Because the server has already validated the client, this is effectively a courtesy offered, and has no impact on authorization to access the protected resources.

\subsubsection{Connection Latency Reporting May Vary}
This work offers a method to deal with latency mismatches between the server and the client by rounding to the nearest prime based latency window. Multiple COTPs should be done through the client, or, simply retrying authentication as proposed in this work. However, this method means that a stochastic method to validating the user should be used to reduce the number of retries. This is in contrast to traditional one-time password methods, which can be performed from a separate device.

\subsection{Future Work}
Future work includes further addressing timing discrepancies between server and client. One interesting approach could be remap the values to nearest prime number or calibrating the error out during server boot. Care must be taken to ensure that close values map to a similar value while values with large differences map to entirely different values - all without the counterpart's number. If such a technique is found, this factor becomes significantly more robust to one time code differences.

Another potential future work is finding the optimal balance between connection oriented details and other factors. This work focused on introducing connection oriented features to a verification algorithm. This is by using only shared features between both parties, however there are additional features that are shared, which can be included. One such additional feature is each party's supported cipher suite, or, another is the selected curve - and both could increase difficultly for an advanced transparent MITM to successfully mimmic all details symmetrically. However, the increase in difficultly also needs to be assessed. Another issue is balance between authentication features, ie., a completely connection-oriented approach will fail, due to connection details being publicly available to a motivated attacker through protocol analysis.

Finally, another future work is potientialy repeatedly validating or checking factors periodically. TLS addresses forward secrecy through rekeying, and this happens to be part of the protocol. The same process could trigger a transparent COTP generation and exchange. This type of feature is already possible on the browser side, and could be implemented on the server side as well. In addition, a new shared secret could be generated and shared upon rekeying, forward authentication is allowed. Those with access to previous details would not have future authentication.

@@
